---
title: "Project 1"
output: html_document
---

**Introduction**
  
League of Legends, widely known simply as LoL, is a PC game categorized as a "MOBA," or "Multiplayer Online Battle Arena."  It is a game in which two teams of 5 players each face off to achieve the ultimate goal of destroying the opposing team's Nexus.  Each player controls a single "champion" who gets stronger as the game progresses, buying items with gold.  Gold is gained by killing "minions" that continuously push towards each enemy's nexus, taking down neutral monsters or enemy structures, and killing enemy champions.  All champions also receive a set amount of gold continuously throughout the game called "ambient gold."  Generally, a team with a gold lead has an advantage in fighting strength that becomes more significant as the gap in gold increases.  A sizable gold lead thus bestows a much higher chance of victory.  

To clarify some of the statistics tracked in LoL, there are 5 that will be studied in this analysis: win rate, KDA ratio, gold per minute, minions killed, and champions killed.  Win rate is self-explanatory.  KDA ratio refers to (K+A)/D, in which K is the number of kills, A is the number of assists, and D is the number of deaths.  Gold per minute is the average amount of gold earned per minute of a game, taking into account any sources (e.g., minion kills, champion kills, ambient gold, etc.).  Finally, minions and champions killed refer to how many kills were achieved against enemy minions and champions.  There is far more to the game that has not been explained, but for the purpose of understanding this report, the explanation is sufficient.

The rise in LoL's popularity as the most popular game in the world has been accompanied by the explosive growth of its professional scene.  In North America and Europe, it is called the League of Legends Championship Series (LCS), of which there is both an NA and EU variant.  A single season of LCS consists of 8 teams and 28 total games.  Each team plays every other team in its respective region 4 times and is ranked at the end of the split by its number of wins.  The data that I will be analyzing is based on the combined stats of NA and EU's 2014 summer split.

The question to be addressed is: Is there a correlation between win rate and KDA ratio, gold per minute, minions killed, or champions killed, in games of LoL at the professional level?  In reality, there are many different ways to find victory in LoL.  Depending on what champions are picked and how each team plays, these can vary from chaotic 5v5 team fights, picking off singled out opponents, outmaneuvering the enemy in taking down structures, and everything in between.  In addition, statistics such as the number of enemies killed or gold acquired are far more valuable when studied relative to the enemy.  For example, having a vast amount of gold means little if the opponent has just as much.  It is thus unclear as to whether there are any variables that are consistently correlated with win rate.  Revealing any variables that show such correlations could provide evidence for the most effective and ineffective strategies across various playstyles, which is a topic that has been debated in the game for years.  Professional players have always struggled to find the best strategies; whether any are objectively (or at least evidently) superior to the rest remains to be discovered.

I would like to begin this analysis of win rate correlation by studying the aforementioned variables: KDA ratio, gold per minute, minions killed, and champions killed.  First, I will draw a scatter plot to check the distribution.  If the data seems to follow a normal distribution, I will test for parametric Pearson correlation to determine correlation between each of the independent variables and win rate.  Otherwise, I will test for non-parametric Spearman correlation.  I hypothesize that each of these variables are positively correlated with win rate due to the fact that increasing any of these variables either grants or is indicative of a team's advantage in game.  I believe this advantage will more often than not be translated into victory.

**Data**

```{r,echo=FALSE}
LCS.table<-read.table("LCS Statistics (Summer 2014).txt",sep="|",header=TRUE,stringsAsFactor=FALSE)
LCS.table
```

The above table shows some of the team-specific statistics that are published on the official League of legends e-sports website for the 2014 summer split (http://na.lolesports.com/na-lcs/2014/split2/).  From left to right, the table conveys: standing of the team in its region, region of LCS, team name, average KDA ratio per game, average gold per minute per game, average minions killed per game, average champions killed per game, total gold accumulated in the entirety of the split in millions, total wins, and total losses.  All 16 teams in NA and EU LCS are accounted for.

**Methods**

```{r,echo=FALSE}
LCS.table$Win.Rate<-((LCS.table$Wins)/28) #Adds extra column calculating win rate of each team
LCS.table
```

The above table was modified to add an extra column showing the calculated win rate for each team from the number of wins and total number of games.  This will be necessary for statistical calculations.

```{r,echo=FALSE}
#Draws scatter plots for each of the variables vs. win rate

plot(LCS.table$Win.Rate,LCS.table$KDA.Ratio,col="blue",pch=16,xlab="Win Rate",ylab="KDA Ratio",main="Average KDA Ratio")
plot(LCS.table$Win.Rate,LCS.table$Gold.Min,col="red",pch=16,xlab="Win Rate",ylab="Gold per Minute",main="Average Amount of Gold Gained per Minute")
plot(LCS.table$Win.Rate,LCS.table$Minions.Killed,col="green",pch=16,xlab="Win Rate",ylab="Minions Killed",main="Average Number of Minions Killed")
plot(LCS.table$Win.Rate,LCS.table$Champions.Killed,col="orange",pch=16,xlab="Win Rate",ylab="Champions Killed",main="Average Number of Champions Killed")
```

The above scatter plots compare each of the variables of interest to win rate.  Judging by the shape of the distributions, it seems reasonable to assume that the relatively linear distributions of the "Average KDA Ratio"" and "Average Amount of Gold Gained per Minute" graphs indicate that the data from those variables follow a relatively normal distribution.  Thus, I will test for Pearson correlation for these variables.  In contrast, the distributions of the Average Number of Minions Killed" and "Average Number of Champions Killed" graphs seem to lack a uniform distribution.  In these cases, I will test for Spearman correlation.

```{r,echo=FALSE}
KDA.Ratio.cor<-cor.test(LCS.table$Win.Rate,LCS.table$KDA.Ratio,method="pearson") #Pearson correlation for KDA ratio
KDA.Ratio.cor
Gold.Min.cor<-cor.test(LCS.table$Win.Rate,LCS.table$Gold.Min,method="pearson") #Pearson correlation for gold per minute
Gold.Min.cor
Minion.Killed.cor<-cor.test(LCS.table$Win.Rate,LCS.table$Minions.Killed,method="spearman") #Spearman correlation for minions killed
Minion.Killed.cor
Champions.Killed.cor<-cor.test(LCS.table$Win.Rate,LCS.table$Champions.Killed,method="spearman") #Spearman correlation for champions killed
Champions.Killed.cor
```

**Results**

```{r,echo=FALSE}
#Displays the relevant results of correlation tests in a data fram
LCS.cor<-data.frame(Variable=c("KDA.Ratio","Gold.Min","Minions.Killed","Champions.Killed"),Test=c("Pearson","Pearson","Spearman","Spearman"),Correlation.Coefficient=c(0.8461,0.9625,0.3044,0.7008),p.Value=c(3.619e-05,2.532e-09,0.2517,0.002492),Signficant=c("Yes","Yes",'No','Yes'))
LCS.cor
```

As observed in the above table, KDA ratio, gold per minute, and champions killed all had correlation coefficients near 1 after applying Spearman and Pearson correlation tests, indicating strong positive correlations.  In addition, each of the p-values for those 3 results were well under the 0.05 threshold, making them statistically significant.

However, the minions killed has a high p-value and relatively low correlation coefficient.  This shows that even if the p-value marked the result as significant (which it did not), there would only be a slight positive correlation.

**Discussion**

As shown in the results, all of the variables are positively correlated with win rate except for minions killed.  That is, statistically speaking, as each of the positively correlated variables increases, the win rate tends to increase along with it.  The number of minions killed, however, does not seem to reflect any statistically significant correlation.  The average number of minions killed per game was hypothesized to show a positive correlation with win rate due to the winning team theoretically having more gold with more minion kills by the end of a given game.  However, the data does not support this hypothesis.  The other variables were also hypothesized with the same assumption in mind: higher KDA, champions killed, and gold per minute should give the team with the higher of these statistics the greater amount of gold by the end of a given game.  But in contrast, these other statistics align with the hypothesis.

A high KDA ratio is indicative of a team’s high number of champion kills and assists relative to its number of champion deaths.  A high number of kills and assists gives a team more gold, while a low number of deaths indicates that a team suffered few deaths to the enemy team and thus prevented the enemy team from gaining much gold through champion kills.  Along with giving a team additional gold, the assassination of an enemy champion also confers other advantages while the enemy champion waits to re-spawn, including free reign to kill minions without interference and greater control of the map (playing field).  These are advantages that give a team a greater chance at victory, possibly affecting the KDA ratio’s positive correlation with win rate.  The relatively low number of deaths also denies the enemy team the same advantage.  It is therefore understandable that a high KDA ratio associated with a high win rate.

The average number of champions killed is a similar statistic to KDA ratio except that it only takes into account the total number of enemy champions killed without considering deaths to the enemy team or assists.  By only studying the number of kills in a match, much less can be inferred because nothing is known about the performance of the enemy team.  Thus, the only statement that can confidently be made from these results is that champion kills are positively correlated with win rate; killing more enemy champions is associated with a higher win rate.

Gold per minute is, like champion kills, a statistic that is limited by lack of more specific data in making assertions beyond the observation that it is positively correlated with win rate.  Average gold per minute in a match can be very volatile; how much gold a team is gaining can vary significantly throughout a single match, and the enemy’s average gold per minute is also unknown.  But it can at least be asserted that the higher a team’s gold per minute is, the higher its win rate is.

It is also important to note that the calculated values are not necessarily as accurate as they could be due to the small sample size.  There were only a total of 16 teams in the LCS 2014 summer split, which is far from an ideal number of samples.  Also, two of the tests yieled a warning that p-value calculation may not be exact due to "ties," or identical data.  The impact should be minimal, but it is still worth noting.

**Future Directions**

The results of analyzing the data yield some very general insight into pro League of Legends strategy.  With that said, it is unfortunately too vague to be of too much use.  Further analysis is necessary to better understand what factors are most commonly correlated with victory.  This would require gathering data that is much more specific, involving match-by-match observation for specific times, differences in gold at those times, champion selection, etc.  Although, this would actually boost the accuracy of the statistical tests due to the larger sample sizes.  Instead of analyzing the averages of each team, analyzing individual matches allows for a sample size equal to all the matches played during the split (>100).  But if I were to name a specific direction that most interests me, it would be in exploration of gold differences.  My next step would be to mark gold differences between teams at pre-determined time points in each match.  The resulting data would be invaluable for answering questions such as: "How much of a gold lead at a certain time in a match gives a statistically significant win rate of above (some arbitrary percentage)?"  The results would be far more applicable in a game where numerous factors prevent too simplistic of an analytic approach from being very effective.
